---
title: "GSEA endothelial cells"
author: "Mechthild Lütge"
date: "10 Dec 2021"
output:
  html_document:
    self_contained: no
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
set.seed(1000)
```

## load packages
```{r libs}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(tidyverse)
  library(Seurat)
  library(magrittr)
  library(dplyr)
  library(purrr)
  library(ggplot2)
  library(here)
  library(runSeurat3)
  library(ggsci)
  library(pheatmap)
  library(ggpubr)
  library(biomaRt)
  library(org.Mm.eg.db)
})

```

## set dir
```{r set dir}
basedir <- here()
seurat <- readRDS(file = paste0(basedir,
                      "/data/MuHeart_totalCellsPlusTreated_integratedLabeled", 
                      "_seurat.rds"))
Idents(seurat) <- seurat$label

```



## color vectors
```{r colVec}

colPal <- pal_igv()(length(levels(seurat)))
colGrp <- pal_jama()(length(unique(seurat$grp)))
colAge <- pal_uchicago()(length(unique(seurat$age)))
colStype <- pal_aaas()(length(unique(seurat$seqType)))
colDat <- c(pal_npg()(10),
            pal_futurama()(10))[1:(length(unique(seurat$dataset)))]
colLab <- pal_igv()(length(unique(seurat$label)))
colTreat <- pal_jama()(length(unique(seurat$treatment)))

  
names(colPal) <- levels(seurat)
names(colGrp) <- unique(seurat$grp)
names(colAge) <- unique(seurat$age)
names(colStype) <- unique(seurat$seqType)
names(colDat) <- unique(seurat$dataset)
names(colLab) <- unique(seurat$label)
names(colTreat) <- unique(seurat$treatment)
```


## vis data {.tabset}
### clusters
```{r vis cluster}

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```

### group
```{r vis grp}

DimPlot(seurat, reduction = "umap", group.by = "grp", cols=colGrp)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

```


### Sample
```{r vis sample}

DimPlot(seurat, reduction = "umap", group.by = "dataset", cols=colDat)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```


### Age
```{r vis age}

DimPlot(seurat, reduction = "umap", group.by = "age", cols=colAge)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```


### SeqType
```{r vis seqType}

DimPlot(seurat, reduction = "umap", group.by = "seqType", cols=colStype)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```

### Treatment
```{r vis treat}

DimPlot(seurat, reduction = "umap", group.by = "treatment", cols=colTreat)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```


## subset and create sce object

```{r sce object}
seuratSub <- subset(seurat, label %in% c("Tcells", "Macrophages", "InfMonos",
                                         "Fibroblasts"))


seuratSub$label_plus_grp_plus_treat <- paste0(seuratSub$label, "_",
                                              seuratSub$grp, "_",
                                              seuratSub$treatment)
seuratSub$grp_plus_treat <- paste0(seuratSub$grp, "_", seuratSub$treatment)

table(seuratSub$label_plus_grp_plus_treat)
downSam <- min(table(seuratSub$label_plus_grp_plus_treat))
downSam <- 500
print(downSam)

Idents(seuratSub) <- seuratSub$label_plus_grp_plus_treat
seuratSub <- subset(seuratSub, downsample = downSam)

sceMerge <- as.SingleCellExperiment(seuratSub)
(table(sceMerge@colData$label_plus_grp_plus_treat))

```


# create count matrix for cellphonedb
```{r create cnt mat}

## set human ensIDs as rownames
ensIDs <- gsub("\\..*","",rownames(sceMerge))

## map to human orthologs
mart2 <- useMart("ensembl", dataset="hsapiens_gene_ensembl")
mart1 = useMart("ensembl", dataset="mmusculus_gene_ensembl") 

# human / mouse
geneIDS_hum <- getLDS(attributes=c("ensembl_gene_id"),
           filters="ensembl_gene_id", values=ensIDs, mart=mart1,
           attributesL=c("ensembl_gene_id"), martL=mart2)

geneIDsMatch <- as.data.frame(ensIDs) %>% 
  left_join(., geneIDS_hum, by=c("ensIDs"="Gene.stable.ID")) 
geneIDsMatch$Gene.stable.ID.1[
  which(is.na(geneIDsMatch$Gene.stable.ID.1))] <- "unknown"
geneIDsMatch <- geneIDsMatch[!duplicated(geneIDsMatch$ensIDs),]
rowData(sceMerge)$humGene <- geneIDsMatch$Gene.stable.ID.1
dim(sceMerge)
sceMerge <- sceMerge[which(rowData(sceMerge)$humGene != "unknown"),] 
dim(sceMerge)

## filter for grp and write cnt mat plus metadat
lapply(unique(sceMerge$grp_plus_treat), function(grp){
  sceGrp <- sceMerge[,which(sceMerge@colData$grp_plus_treat == grp)]
  cellPhoneGrp <- cbind(rowData(sceGrp)$humGene, as.matrix(logcounts(sceGrp)))
  colnames(cellPhoneGrp)[1] <- "gene"
  write.table(cellPhoneGrp, 
              file = paste0(basedir,
                            "/data/cellphonedb/totalCellsSubsampled_",grp,
                            "_countMatrix.txt"),
              row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")

  metaDatGrp <- as.data.frame(cbind(colnames(sceGrp),
                                    colData(sceGrp)$label))
  colnames(metaDatGrp) <- c("Cell", "cell_type")
  write.table(metaDatGrp, 
              file = paste0(basedir,
                            "/data/cellphonedb/totalCellsSubsampled_",grp,
                            "_metaData.txt"),
              row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")


})



```


## session info
```{r session info}
sessionInfo()
date()
```


cd /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb
eval "$(pyenv init -)"
eval "$(pyenv virtualenv-init -)"

## run cellphonedb
### build database
cellphonedb database generate --user-interactions custom_interaction_file.csv

##run statistical analysis
## subsampling in R script.. 
cellphonedb method statistical_analysis /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/totalCellsSubsampled_TCRM_no_metaData.txt /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/totalCellsSubsampled_TCRM_no_countMatrix.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/outTCRM_no/ --database /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb/out/cellphonedb_user_2021-10-05-09_17.db --debug-seed 3 --iterations 100

cellphonedb method statistical_analysis /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/totalCellsSubsampled_control_no_metaData.txt /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/totalCellsSubsampled_control_no_countMatrix.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/outControl_no/ --database /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb/out/cellphonedb_user_2021-10-05-09_17.db --debug-seed 3 --iterations 100

cellphonedb method statistical_analysis /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/totalCellsSubsampled_TCRM_isotype_metaData.txt /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/totalCellsSubsampled_TCRM_isotype_countMatrix.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/outTCRM_isotype/ --database /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb/out/cellphonedb_user_2021-10-05-09_17.db --debug-seed 3 --iterations 100

cellphonedb method statistical_analysis /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/totalCellsSubsampled_TCRM_14D102_metaData.txt /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/totalCellsSubsampled_TCRM_14D102_countMatrix.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/outTCRM_14D102/ --database /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb/out/cellphonedb_user_2021-10-05-09_17.db --debug-seed 3 --iterations 100


Count network interactions:
cellphonedb plot heatmap_plot /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/totalCellsSubsampled_TCRM_no_metaData.txt --pvalues-path /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/outTCRM_no/pvalues.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/outTCRM_no

cellphonedb plot heatmap_plot /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/totalCellsSubsampled_control_no_metaData.txt --pvalues-path /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/outControl_no/pvalues.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/outControl_no

cellphonedb plot heatmap_plot /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/totalCellsSubsampled_TCRM_isotype_metaData.txt --pvalues-path /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/outTCRM_isotype/pvalues.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/outTCRM_isotype

cellphonedb plot heatmap_plot /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/totalCellsSubsampled_TCRM_14D102_metaData.txt --pvalues-path /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/outTCRM_14D102/pvalues.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/TCRM_Christina/data/cellphonedb/outTCRM_14D102

We use CellPhoneDB as tool to predict cellular communication in our dataset. In brief, pairs of known interacting ligands and receptors are derived from public databases and cell types are analysed for enriched receptor–ligand interactions based on expression of a receptor by one cell type and a ligand by another cell type. 

deactivate  


