---
title: "Merge samples"
author: "Mechthild Lütge"
date: "14 May 2020"
output:
  html_document:
    self_contained: no
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1000)
```

## load packages
```{r libs}
suppressPackageStartupMessages({
  library(tidyverse)
  library(Seurat)
  library(magrittr)
  library(dplyr)
  library(purrr)
  library(ggplot2)
  library(here)
  library(runSeurat3)
  library(ggsci)
  library(pheatmap)
})

```

## set dir
```{r set dir}
basedir <- here()
metaDat <- read_tsv(paste0(basedir, "/metadata.txt"), col_names = T)

```


## load and assign samples
```{r load plus assign samples}

assignSamples <- function(smpNam, basedirSmp, smpGrp, smpBatch){
  smpNamFull <- list.files(path = paste0(basedirSmp, "/data/"),
				 pattern = paste0(smpNam, ".*_seurat.rds"))
  seuratSmp <- readRDS(paste0(basedirSmp, "/data/", smpNamFull))
  seuratSmp$grp <- smpGrp
  seuratSmp$batch <- smpBatch
  return(seuratSmp)
}

####################################################################

for(i in 1:length(metaDat$Sample)){
  seuratX <- assignSamples(smpNam = metaDat$Sample[i],
                           basedirSmp = basedir,
                           smpGrp = metaDat$group[i],
                           smpBatch = metaDat$batch[i])
  if(exists("seurat")){
    seurat <- merge(x = seurat, y = seuratX, project = "TCRMxCcl19")
  }else{
    seurat <- seuratX
  }
}

remove(seuratX)

```



## run clustering and DR and remove contaminating cells
```{r clustering and DR}
seurat <- rerunSeurat3(seurat)

## remove cluster 9 (cardiomyocytes: Fabp3, Tnnc1, Tnnt2, Myl2, Myl3, Myh6)
seurat <- subset(seurat, idents=9, invert=T)

## remove cluster 10 (epithelial cells + endothelial cells: Krt19, Cldn5, Cdh5,
## Esam, Pecam1)
seurat <- subset(seurat, idents=10, invert=T)

## remove cluster 11 (proliferating cells: Mki67, Cdk1, Ckap2
seurat <- subset(seurat, idents=11, invert=T)

## rerun DR
seurat <- RunPCA(object = seurat, npcs = 30, verbose = FALSE)
seurat <- RunTSNE(object = seurat, reduction = "pca", dims = 1:20)
seurat <- RunUMAP(object = seurat, reduction = "pca", dims = 1:20)


dat <- data.frame(table(seurat$dataset))
knitr::kable(dat)
```
## color vectors
```{r colVec}

colPal <- pal_igv()(length(levels(seurat)))
colGrp <- pal_jama()(length(unique(seurat$grp)))
colSmp <- pal_uchicago()(length(unique(seurat$dataset)))

names(colPal) <- levels(seurat)
names(colGrp) <- unique(seurat$grp)
names(colSmp) <- unique(seurat$dataset)
```


## vis data {.tabset}
### clusters
```{r vis cluster}
DimPlot(seurat, reduction = "umap", cols=colPal)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```

### group
```{r vis grp}

DimPlot(seurat, reduction = "umap", group.by = "grp", cols=colGrp)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

```


### Sample
```{r vis sample}

DimPlot(seurat, reduction = "umap", group.by = "dataset", cols=colSmp)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```


## marker genes
```{r marker genes}

seurat_markers_all <- FindAllMarkers(object = seurat, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

```

### top 15 marker genes per cluster {.tabset}

```{r avg heatmap function, eval=T, include=FALSE}

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  ## format gene names (depends on how gene list is inputed)
  selGenes <- selGenes %>% 
      mutate(labelNam = (str_split(gene, '\\.', simplify = T)[,2]))
  selGenes <- selGenes$labelNam
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      cond = colVecCond,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

```


```{r top marker}

cluster <- levels(seurat)
selGenesAll <- seurat_markers_all %>% group_by(cluster) %>% 
  top_n(-15, p_val_adj) %>% 
  top_n(15, avg_logFC)

template_hm <- c(
    "#### {{cl}}\n",
    "```{r top marker {{cl}}, fig.height=8, fig.width=6, echo = FALSE}\n",
    "selGenes <- selGenesAll %>% filter(cluster=='{{cl}}')",
    "pOut <- avgHeatmap(seurat = seurat, selGenes = selGenes,
                  colVecIdent = colPal, 
                  ordVec=levels(seurat),
                  gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                  cr=T, condCol=F)\n",
    "```\n",
    "\n"
  )

plots_gp <- lapply(cluster, 
  function(cl) knitr::knit_expand(text = template_hm)
)

```

`r knitr::knit(text = unlist(plots_gp))`

## overall DE genes
```{r overall DE, fig.height=10, fig.width=6}

Idents(seurat) <- seurat$grp

DEgenes <- FindAllMarkers(object = seurat, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

### plot DE genes top 15 avg logFC
DEGenesAll <- DEgenes %>% group_by(cluster) %>% 
  top_n(15, avg_logFC)

Idents(seurat) <- seurat$RNA_snn_res.0.4
avgHeatmap(seurat = seurat, selGenes = DEGenesAll,
                  colVecIdent = colPal, 
                  ordVec=levels(seurat),
                  gapVecR=NULL, gapVecC=NULL,cc=T,
                  cr=T, condCol=F)
```


## save objects
```{r save}
saveRDS(seurat, file = paste0(basedir, "/data/TCRM_allMerged_seurat.rds"))

write.table(seurat_markers_all,
            file=paste0(basedir, "/data/TCRM_allMerged_markerGenes.txt"),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")

write.table(DEgenes,
            file=paste0(basedir, "/data/TCRM_allMerged_DEGenes.txt"),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")

```


## TCRM only
```{r TCRM only}
(table(seurat$grp, seurat$seurat_clusters))
## for characterization we subset on clusters pre-dominantly present in TCRM  
seuratSub <- subset(seurat, idents = c("0", "2", "3"), invert=T)
(table(seuratSub$grp, seuratSub$seurat_clusters))
```

### vis data {.tabset}
#### clusters
```{r vis cluster TCRM}
DimPlot(seuratSub, reduction = "umap", cols=colPal)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```

#### Sample
```{r vis sample TCRM}

DimPlot(seuratSub, reduction = "umap", group.by = "dataset", cols=colSmp)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```


### marker genes
```{r marker genes TCRM}
seurat_markers_all <- FindAllMarkers(object = seuratSub, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

```

### top 15 marker genes per cluster {.tabset}

```{r top marker TCRM}

cluster <- levels(seuratSub)
selGenesAll <- seurat_markers_all %>% group_by(cluster) %>% 
  top_n(-15, p_val_adj) %>% 
  top_n(15, avg_logFC)

template_hm <- c(
    "#### {{cl}}\n",
    "```{r top marker TCRM {{cl}}, fig.height=8, fig.width=6, echo = FALSE}\n",
    "selGenes <- selGenesAll %>% filter(cluster=='{{cl}}')",
    "pOut <- avgHeatmap(seurat = seuratSub, selGenes = selGenes,
                  colVecIdent = colPal, 
                  ordVec=levels(seuratSub),
                  gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                  cr=T, condCol=F)\n",
    "```\n",
    "\n"
  )

plots_gp <- lapply(cluster, 
  function(cl) knitr::knit_expand(text = template_hm)
)

```

`r knitr::knit(text = unlist(plots_gp))`

## save objects
```{r save TCRM}
saveRDS(seuratSub, file = paste0(basedir, "/data/TCRM_TCRM_seurat.rds"))

write.table(seurat_markers_all,
            file=paste0(basedir, "/data/TCRM_TCRM_markerGenes.txt"),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
```


## Control only
```{r Control only}
(table(seurat$grp, seurat$seurat_clusters))
## for characterization we subset on clusters pre-dominantly present in TCRM  
seuratSub <- subset(seurat, idents = c("0", "2", "3", "7"))
(table(seuratSub$grp, seuratSub$seurat_clusters))
```

### vis data {.tabset}
#### clusters
```{r vis cluster Control}
DimPlot(seuratSub, reduction = "umap", cols=colPal)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```

#### Sample
```{r vis sample Control}

DimPlot(seuratSub, reduction = "umap", group.by = "dataset", cols=colSmp)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```


### marker genes
```{r marker genes Control}
seurat_markers_all <- FindAllMarkers(object = seuratSub, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

```

### top 15 marker genes per cluster {.tabset}

```{r top marker Control}

cluster <- levels(seuratSub)
selGenesAll <- seurat_markers_all %>% group_by(cluster) %>% 
  top_n(-15, p_val_adj) %>% 
  top_n(15, avg_logFC)

template_hm <- c(
    "#### {{cl}}\n",
    "```{r top marker Control {{cl}}, fig.height=8,fig.width=6,echo = FALSE}\n",
    "selGenes <- selGenesAll %>% filter(cluster=='{{cl}}')",
    "pOut <- avgHeatmap(seurat = seuratSub, selGenes = selGenes,
                  colVecIdent = colPal, 
                  ordVec=levels(seuratSub),
                  gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                  cr=T, condCol=F)\n",
    "```\n",
    "\n"
  )

plots_gp <- lapply(cluster, 
  function(cl) knitr::knit_expand(text = template_hm)
)

```

`r knitr::knit(text = unlist(plots_gp))`

## save objects
```{r save Control}
saveRDS(seuratSub, file = paste0(basedir, "/data/TCRM_Control_seurat.rds"))

write.table(seurat_markers_all,
            file=paste0(basedir, "/data/TCRM_Control_markerGenes.txt"),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
```




## session info
```{r session info}
sessionInfo()
date()
```




