---
title: "Merge CD45+ cells"
author: "Mechthild Lütge"
date: "14 May 2020"
output:
  html_document:
    self_contained: no
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1000)
```

## load packages
```{r libs}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(tidyverse)
  library(Seurat)
  library(magrittr)
  library(dplyr)
  library(purrr)
  library(ggplot2)
  library(here)
  library(runSeurat3)
  library(ggsci)
  library(ggpubr)
  library(pheatmap)
})

```

## set dir
```{r set dir}
basedir <- here()
metaDat <- read_tsv(paste0(basedir, "/metadata_lymphcytes.txt"), col_names = T)

```


## load and assign samples
```{r load plus assign samples}

assignSamples <- function(smpNam, basedirSmp, smpGrp, smpBatch){
  smpNamFull <- list.files(path = paste0(basedirSmp, "/data/"),
				 pattern = paste0(smpNam, ".*_seurat.rds"))
  seuratSmp <- readRDS(paste0(basedirSmp, "/data/", smpNamFull))
  seuratSmp$grp <- smpGrp
  seuratSmp$batch <- smpBatch
  return(seuratSmp)
}

####################################################################

for(i in 1:length(metaDat$Sample)){
  seuratX <- assignSamples(smpNam = metaDat$Sample[i],
                           basedirSmp = basedir,
                           smpGrp = metaDat$group[i],
                           smpBatch = metaDat$batch[i])
  if(exists("seurat")){
    seurat <- merge(x = seurat, y = seuratX, project = "TCRMxCD45")
  }else{
    seurat <- seuratX
  }
}

remove(seuratX)

```


## remove contaminating cells
```{r remove cont cells}

seurat <- subset(seurat, subset = ENSMUSG00000001506.Col1a1 == 0)
seurat <- subset(seurat, subset = ENSMUSG00000029231.Pdgfra == 0)
seurat <- subset(seurat, subset = ENSMUSG00000024620.Pdgfrb == 0)
seurat <- subset(seurat, subset = ENSMUSG00000020717.Pecam1 == 0)

```


## run clustering and DR and remove contaminating cells
```{r clustering and DR}

seurat <- rerunSeurat3(seurat, res=c(0.6, 0.4, 0.25))
dim(seurat)

```

## color vectors
```{r colVec}

colPal <- pal_igv()(length(levels(seurat)))
colGrp <- pal_jama()(length(unique(seurat$grp)))
colSmp <- pal_uchicago()(length(unique(seurat$dataset)))

names(colPal) <- levels(seurat)
names(colGrp) <- unique(seurat$grp)
names(colSmp) <- unique(seurat$dataset)
```

## vis data {.tabset}
### clusters
```{r vis cluster}
DimPlot(seurat, reduction = "umap", cols=colPal)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```

### group
```{r vis grp}

DimPlot(seurat, reduction = "umap", group.by = "grp", cols=colGrp)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

```

## marker genes
```{r marker genes}

seurat_markers_all <- FindAllMarkers(object = seurat, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

```


```{r avg heatmap function, eval=T, include=FALSE}

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(label=gsub("^.*\\.", "", gene))
  ## format gene names (depends on how gene list is inputed)
  selGenes <- selGenes %>% 
      mutate(labelNam = (str_split(gene, '\\.', simplify = T)[,2])) %>% 
    filter(labelNam %in% genes$label)
  #selGenes <- selGenes$labelNam
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  #genes <- data.frame(gene=rownames(seurat)) %>% 
  #  mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% selGenes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  #genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      cond = colVecCond,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes$labelNam,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

```


## overall DE genes
```{r overall DE, fig.height=8, fig.width=6}

Idents(seurat) <- seurat$grp

DEgenes <- FindAllMarkers(object = seurat, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

### plot DE genes top 15 avg logFC
DEGenesAll <- DEgenes %>% group_by(cluster) %>% 
  top_n(15, avg_logFC)

Idents(seurat) <- seurat$RNA_snn_res.0.25
avgHeatmap(seurat = seurat, selGenes = DEGenesAll,
                  colVecIdent = colPal, 
                  ordVec=levels(seurat),
                  gapVecR=NULL, gapVecC=NULL,cc=T,
                  cr=T, condCol=F)
```



## clusterwise DE genes

```{r clusterwise DE genes}

seurat$clustMerge <- seurat$RNA_snn_res.0.25
Idents(seurat) <- seurat$grp
grpVec <- levels(seurat$clustMerge)

clustDE <- lapply(grpVec, function(grp){
  grpSub <- levels(seurat$clustMerge)[which(levels(seurat$clustMerge)==grp)]
  seuratSub <- subset(seurat, clustMerge == grpSub)
  DEgenes <-FindAllMarkers(seuratSub, only.pos=T, logfc.threshold = 0.25,
                           min.pct = 0.1)
  DEgenes <- DEgenes %>% filter(p_val_adj<0.01) %>%
    mutate(group=paste0(grp, "_", cluster))
})

names(clustDE) <- grpVec

clustDE_Dat <- data.frame(do.call("rbind", clustDE))
```



### avg Heatmap top cwDE genes {.tabset}
```{r avgHeat cwDE, warning=FALSE, echo = FALSE}

## only cluster with > 1 DE genes
deCnt <- lapply(clustDE, nrow)
deCnt <- data.frame(cnt=do.call("rbind", deCnt)) %>% rownames_to_column(var = "cluster") %>% filter(cnt>1)

grpVec <- as.character(deCnt$cluster)
seurat$cluster_plus_conds <- as.factor(paste0(seurat$clustMerge,
                                              "_",seurat$grp))
Idents(seurat) <- seurat$cluster_plus_conds

template_ah <- c(
    "#### {{grp}}\n",
    "```{r avgHeat cwDE {{grp}}, echo = FALSE, fig.height=10, fig.width=10}\n",
    "DeSub <- clustDE[[as.character('{{grp}}')]]",
    "DeSel <- DeSub %>% top_n(., 30, avg_logFC)",
    "gapVecCol <- seq(2, length(levels(seurat$cluster_plus_conds)), by=2)",
    "pOut <- avgHeatmap(seurat = seurat, selGenes = DeSel,
                     colVecIdent = colPal, colVecCond=colGrp,
                     ordVec=levels(seurat$cluster_plus_conds),
                     gapVecR=NULL, gapVecC=gapVecCol,cc=FALSE,
                     cr=T, condCol=T)\n",
    "```\n",
    "\n"
)
    
plots_ah <- lapply(grpVec, 
  function(grp) knitr::knit_expand(text = template_ah)
)

```

`r knitr::knit(text = unlist(plots_ah))`


## save objects
```{r save}
Idents(seurat) <- seurat$seurat_clusters
saveRDS(seurat, file = paste0(basedir, 
                              "/data/Mu_Heart_TCRMplusLM_CD45_seurat.rds"))

write.table(seurat_markers_all,
            file=paste0(basedir, "/data/Mu_Heart_TCRMplusLM_CD45_markerGenes.txt"),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")

write.table(DEgenes,
            file=paste0(basedir,
                        "/data/Mu_Heart_TCRMplusLM_CD45_overall_DEGenes.txt"),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")

write.table(clustDE_Dat,
            file=paste0(basedir,
                        "/data/Mu_Heart_TCRMplusLM_CD45_cwDEGenes.txt"),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")

```


## session info
```{r session info}
sessionInfo()
date()
```




