---
title: "reembedding Fibroblasts"
author: "Mechthild Lütge"
date: "24 July 2020"
output:
  html_document:
    self_contained: no
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
set.seed(1000)
```

## load packages
```{r libs}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(tidyverse)
  library(Seurat)
  library(magrittr)
  library(dplyr)
  library(purrr)
  library(ggplot2)
  library(here)
  library(runSeurat3)
  library(ggsci)
  library(pheatmap)
  library(ggpubr)
})

```

## set dir
```{r set dir}
basedir <- here()
seurat <- readRDS(file = paste0(basedir,
                      "/data/MuHeart_totalCellsPlusTreated_integratedLabeled", 
                      "_seurat.rds"))
Idents(seurat) <- seurat$label
seurat$age[which(seurat$treatment != "no")] <- "12wk"

```



## color vectors
```{r colVec}

colPal <- pal_igv()(length(seurat$integrated_snn_res.0.4))
colGrp <- pal_jama()(length(unique(seurat$grp)))
colAge <- pal_uchicago()(length(unique(seurat$age)))
colStype <- pal_aaas()(length(unique(seurat$seqType)))
colDat <- c(pal_npg()(10),
            pal_futurama()(10))[1:(length(unique(seurat$dataset)))]
colTreat <- pal_jama()(length(unique(seurat$treatment)))

names(colPal) <- unique(seurat$integrated_snn_res.0.4)
names(colGrp) <- unique(seurat$grp)
names(colAge) <- unique(seurat$age)
names(colStype) <- unique(seurat$seqType)
names(colDat) <- unique(seurat$dataset)
names(colTreat) <- unique(seurat$treatment)

colLab <- c("#f4c37e", "#e73213", "#209ab9", "#df7f13",
            "#609eca", "#caa160", "#dbb6ca", "#98bfe8",
            "#7c0606", "#c2a17d", "#db27ab", "#dadada",
            "#32b4a5", "#4c8376")

names(colLab) <- c("InfMonos", "Tcells", "Endothelial", "Macrophages",
                   "Fibroblasts", "pDCsDCs", "NKcells", "Perivascular",
                   "Neutrophils", "Bcells", "Cardiomyocytes", "proliferating",
                   "endocardialEndothelial", "capEndothelial")


```

## reembedding Fibroblasts untreated only
```{r reembedding fbroblasts}
seurat <- subset(seurat, label == "Fibroblasts")
seurat <- subset(seurat, treatment=="no")

### rerun integration
seurat.list <- SplitObject(object = seurat, split.by = "seqType")
for (i in 1:length(x = seurat.list)) {
    seurat.list[[i]] <- NormalizeData(object = seurat.list[[i]],
                                      verbose = FALSE)
    seurat.list[[i]] <- FindVariableFeatures(object = seurat.list[[i]], 
        selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}

seurat.anchors <- FindIntegrationAnchors(object.list = seurat.list, dims = 1:20,
                                         k.filter = 200)
seurat.int <- IntegrateData(anchorset = seurat.anchors, dims = 1:20)
DefaultAssay(object = seurat.int) <- "integrated"

# rerun seurat
seurat.int <- ScaleData(object = seurat.int, verbose = FALSE,
                        features = rownames(seurat.int))
seurat.int <- RunPCA(object = seurat.int, npcs = 20, verbose = FALSE)
seurat.int <- RunTSNE(object = seurat.int, reduction = "pca", dims = 1:20)
seurat.int <- RunUMAP(object = seurat.int, reduction = "pca", dims = 1:20)

seurat.int <- FindNeighbors(object = seurat.int, reduction = "pca", dims = 1:20)
res <- c(0.6,0.8,0.4,0.25)
for(i in 1:length(res)){
  seurat.int <- FindClusters(object = seurat.int, resolution = res[i],
                             random.seed = 1234)
}

DefaultAssay(object = seurat.int) <- "RNA"
seurat <- seurat.int

colPal <- pal_igv()(length(seurat$integrated_snn_res.0.25))
names(colPal) <- unique(seurat$integrated_snn_res.0.25)

```



## vis data all {.tabset}
### clusters
```{r vis cluster all}

DimPlot(seurat, group.by ="integrated_snn_res.0.25", reduction = "umap",
        cols=colPal)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, group.by ="integrated_snn_res.0.25", reduction = "umap",
        cols=colPal)+
  theme_void()
```

### Label
```{r vis label all}

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab,
        shuffle = T)+
  theme_void() 

```

### group
```{r vis grp all}

DimPlot(seurat, reduction = "umap", group.by = "grp", cols=colGrp)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "grp", cols=colGrp, shuffle = T)+
  theme_void() 

```


### Sample
```{r vis sample all, fig.height=4, fig.width=7}

DimPlot(seurat, reduction = "umap", group.by = "dataset", cols=colDat)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "dataset", cols=colDat)+
  theme_void()

```


### Age
```{r vis age all}

DimPlot(seurat, reduction = "umap", group.by = "age", cols=colAge)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "age", cols=colAge, shuffle = T)+
  theme_void() 

```


### SeqType
```{r vis seqType all}

DimPlot(seurat, reduction = "umap", group.by = "seqType", cols=colStype)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "seqType", cols=colStype)+
  theme_void() 
```

### Treatment
```{r vis treat all}

DimPlot(seurat, reduction = "umap", group.by = "treatment", cols=colTreat)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "treatment", cols=colTreat,
        shuffle = T)+
  theme_void() 

```

### cnt Tab all
```{r cnts par lab per seqType}

table(seurat$label, seurat$seqType)
table(seurat$label, seurat$grp)
```


## expression of selected genes {.tabset}
```{r sel genes}

selGenes <- c("Grem1", "Grem2", "Bmp2","Bmp4", "Bmp7", "Bmpr1a", "Bmpr1b",
              "Bmpr2", "Bst1", "Clu", "Ncam1", "Ccl2", "Cxcl9", "Il11", "Timp1",
              "Plaur", "Hif1a", "Sod2", "Ccl19", "Cxcl13", "Il33", "Il6",
              "Vcam1", "Icam1", "H2-Aa", "Rosa26eyfp", "Ly6a")
selGenesLong <- data.frame(gene=rownames(seurat)) %>%
  mutate(short=gsub("^.*\\.", "", gene)) %>% filter(short %in% selGenes)

```


### sel Genes across all {.tabset}
```{r sel Genes across all}

template_fp <- c(
    "#### {{sel}}\n",
    "```{r sel Genes across all {{sel}},fig.height=3, fig.width=4, echo = FALSE}\n",
    "selGene <- as.character('{{sel}}')",
    "p <- FeaturePlot(seurat, reduction = 'umap',
      features = selGene,
      cols=c('lightgrey', 'darkred'),
              order = T)+
    theme(legend.position='right')",
    "p\n",
    "```\n",
    "\n"
)


plots_fp <- lapply(selGenesLong$gene, 
  function(sel) knitr::knit_expand(text = template_fp)
)

```

`r knitr::knit(text = unlist(plots_fp))`

### sel Genes split by grp {.tabset}
```{r el Genes split by grp}

template_fp <- c(
    "#### {{sel}}\n",
    "```{r sel Genes split by grp {{sel}},fig.height=3, fig.width=8, echo = FALSE}\n",
    "selGene <- as.character('{{sel}}')",
    "p <- FeaturePlot(seurat, reduction = 'umap',
      features = selGene,
      cols=c('lightgrey', 'darkred'),
      order = T,
      split.by = 'grp')+
    theme(legend.position='right')",
    "p\n",
    "```\n",
    "\n"
)


plots_fp <- lapply(selGenesLong$gene, 
  function(sel) knitr::knit_expand(text = template_fp)
)

```

`r knitr::knit(text = unlist(plots_fp))`

### sel Genes split by age {.tabset}
```{r sel Genes split by age}

template_fp <- c(
    "#### {{sel}}\n",
    "```{r sel Genes split by age {{sel}},fig.height=3, fig.width=12, echo = FALSE}\n",
    "selGene <- as.character('{{sel}}')",
    "p <- FeaturePlot(seurat, reduction = 'umap',
      features = selGene,
      cols=c('lightgrey', 'darkred'),
      order = T,
      split.by = 'age')+
    theme(legend.position='right')",
    "p\n",
    "```\n",
    "\n"
)


plots_fp <- lapply(selGenesLong$gene, 
  function(sel) knitr::knit_expand(text = template_fp)
)

```

`r knitr::knit(text = unlist(plots_fp))`


## save object untreated with label
```{r save int untreated}
saveRDS(seurat, file = paste0(basedir,
                      "/data/MuHeart_totalCells_FibroReembedded_wotreat_seurat.rds"))

```


## session info
```{r session info}
sessionInfo()
date()
```




