---
title: "vis BMPs genes TCRM treated versus untreated total cells"
author: "Mechthild Lütge"
date: "24 July 2020"
output:
  html_document:
    self_contained: no
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
set.seed(1000)
```

## load packages
```{r libs}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(tidyverse)
  library(Seurat)
  library(magrittr)
  library(dplyr)
  library(purrr)
  library(ggplot2)
  library(here)
  library(runSeurat3)
  library(ggsci)
  library(pheatmap)
  library(ggpubr)
  library(clusterProfiler)
  library(org.Mm.eg.db)
  library(DOSE)
  library(enrichplot)
  library(viridis)
})

```


## heatmap function

```{r avg heatmap funct}
avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  ## format gene names (depends on how gene list is inputed)
  selGenes <- selGenes %>% 
      mutate(labelNam = (str_split(gene, '\\.', simplify = T)[,2]))
  selGenes <- selGenes$labelNam
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
#    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::mutate(cond = col1) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      cond = colVecCond,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

```



## set dir
```{r set dir}
basedir <- here()
seurat <- readRDS(file = paste0(basedir,
                      "/data/MuHeart_totalCellsPlusTreated_integratedLabeled", 
                      "_seurat.rds"))
Idents(seurat) <- seurat$label
dim(seurat)
seurat <- subset(seurat, treatment == "no", invert=T)
dim(seurat)
```



## color vectors
```{r colVec}

colPal <- pal_igv()(length(levels(seurat)))
colGrp <- pal_jama()(length(unique(seurat$grp)))
colAge <- pal_uchicago()(length(unique(seurat$age)))
colStype <- pal_aaas()(length(unique(seurat$seqType)))
colDat <- c(pal_npg()(10),
            pal_futurama()(10))[1:(length(unique(seurat$dataset)))]
colTreat <- pal_jama()(length(unique(seurat$treatment)))

names(colPal) <- levels(seurat)
names(colGrp) <- unique(seurat$grp)
names(colAge) <- unique(seurat$age)
names(colStype) <- unique(seurat$seqType)
names(colDat) <- unique(seurat$dataset)
names(colTreat) <- unique(seurat$treatment)

colLab <- c("#5050FF", "#CE3D32", "#749B58", "#F0E685",
            "#466983", "#BA6388", "#5DB1DD", "#802268",
            "#6BD76B", "#D595A7", "#924822", "#837B8D",
            "#C75127", "#D58F5C", "#7A65A5", "#E4AF69")

names(colLab) <- c("InfMonos", "Tcells", "Endothelial", "Macrophages",
                   "Fibroblasts", "DCs", "NKcells", "Perivascular",
                   "Neutrophils", "Bcells", "Cardiomyocytes", "proliferating",
                   "pDCs", "endocardialEndothelial", "Eosinophils",
                   "GlialCells")

```


## vis data {.tabset}

### group
```{r vis grp}

DimPlot(seurat, reduction = "umap", group.by = "grp", cols=colGrp)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

```


### Sample
```{r vis sample, fig.height=3.5, fig.width=6}

DimPlot(seurat, reduction = "umap", group.by = "dataset", cols=colDat)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```


### Age
```{r vis age}

DimPlot(seurat, reduction = "umap", group.by = "age", cols=colAge)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```


### SeqType
```{r vis seqType}

DimPlot(seurat, reduction = "umap", group.by = "seqType", cols=colStype)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "seqType", cols=colStype)+
  theme_void() 
```

### Treatment
```{r vis treat}

DimPlot(seurat, reduction = "umap", group.by = "treatment", cols=colTreat)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "treatment", cols=colTreat)+
  theme_void() 

```


### label
```{r vis label}

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab)+
  theme_void() 

```


### label split by seqType
```{r vis label split seq, fig.height=3.5, fig.width=7}

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab,
        split.by = "seqType")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```

### label split by treat
```{r vis label split treat, fig.height=3.5, fig.width=7}

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab,
        split.by = "treatment")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab,
        split.by = "treatment")+
  theme_void() 

```

### col plus split by seqType
```{r vis plus split seqType, fig.height=3.5, fig.width=7}

DimPlot(seurat, reduction = "umap", group.by = "seqType", cols=colStype,
        split.by = "seqType")+
  theme_void() 

```

### col plus split by age
```{r vis plus split age, fig.height=3.5, fig.width=7}

DimPlot(seurat, reduction = "umap", group.by = "age", cols=colAge,
        split.by = "age")+
  theme_void() 

```



## vis BMPs {.tabset}
### violin
```{r vis sel genes violin}

genesDat <- data.frame(EnsID=rownames(seurat)) %>% 
  mutate(gene=gsub(".*\\.", "", EnsID))
selGenes <- data.frame(gene=c("Bmp2", "Bmp4", "Bmpr1a", "Bmpr1b", "Bmpr2",
                              "Bmp7")) %>% 
  left_join(., genesDat, by="gene")

Idents(seurat) <- seurat$treatment
#seuratSub <- subset(seurat, downsample = min(table(seurat$cond2)))

pList <- sapply(selGenes$EnsID, function(x){
  p <- VlnPlot(object = seurat, features = x,
               group.by = "treatment",
               cols = colTreat, pt.size = 0
               )
  plot(p)
})

pList <- sapply(selGenes$EnsID, function(x){
  p <- VlnPlot(object = seurat, features = x,
               group.by = "label", split.by = "treatment",
               cols = colTreat, pt.size = 0
               )
  plot(p)
})


```


### heatmap
```{r vis sel genes heatmap, fig.height=6, fig.width=14}
selGenesHM <- selGenes %>% mutate(gene = EnsID)

ordFix <- c("Bcells_isotype","Bcells_14D102","Tcells_isotype","Tcells_14D102",
            "NKcells_isotype","NKcells_14D102","Eosinophils_isotype",
            "Eosinophils_14D102","InfMonos_isotype","InfMonos_14D102",
            "Macrophages_isotype","Macrophages_14D102","Neutrophils_isotype",
            "Neutrophils_14D102","pDCs_isotype","pDCs_14D102","DCs_isotype",
            "DCs_14D102","proliferating_isotype","proliferating_14D102",
            "Cardiomyocytes_isotype","Cardiomyocytes_14D102",
            "endocardialEndothelial_isotype", "endocardialEndothelial_14D102",
            "Endothelial_isotype","Endothelial_14D102","Fibroblasts_isotype",
            "Fibroblasts_14D102","GlialCells_isotype","GlialCells_14D102",
            "Perivascular_isotype","Perivascular_14D102")

seurat$label_plus_treat <- paste0(seurat$label, "_", seurat$treatment)
seurat$label_plus_treat <- factor(seurat$label_plus_treat, levels = ordFix)
Idents(seurat) <- seurat$label_plus_treat
gapVecCol <- seq(2, length(unique(seurat$label_plus_treat)), by=2)
pOut <- avgHeatmap(seurat = seurat, selGenes = selGenesHM,
                    colVecIdent = colLab, colVecCond=colTreat,
                    ordVec=levels(seurat$label_plus_treat),
                    gapVecR=NULL, gapVecC=gapVecCol,cc=FALSE,
                    cr=T, condCol=T)

```

### dotplot across label
```{r vis sel genes Dotplot across label, fig.height=5, fig.width=9}

Idents(seurat) <- seurat$label_plus_treat
DotPlot(seurat, assay="RNA", features = selGenes$EnsID, scale =T,
        cluster.idents = F, dot.min = 0, dot.scale = 3, scale.by = "size") +
  scale_color_viridis_c() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=selGenes$EnsID, labels=selGenes$gene) +
  xlab("") + ylab("")

DotPlot(seurat, assay="RNA", features = selGenes$EnsID, scale =F,
        cluster.idents = F, dot.min = 0, dot.scale = 5, scale.by = "size") +
  scale_color_viridis_c() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=selGenes$EnsID, labels=selGenes$gene) +
  xlab("") + ylab("")

## only Bmp4 and Bmpr1a
selGenesSub <- selGenes %>% filter(gene %in% c("Bmpr1a", "Bmp4"))
DotPlot(seurat, assay="RNA", features = selGenesSub$EnsID, scale =F,
        cluster.idents = F, dot.min = 0, dot.scale = 5, scale.by = "size") +
  scale_color_viridis_c() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=selGenesSub$EnsID, labels=selGenesSub$gene) +
  xlab("") + ylab("")

```


### dotplot total
```{r vis sel genes Dotplot total, fig.height=4, fig.width=4}

Idents(seurat) <- seurat$treatment
DotPlot(seurat, assay="RNA", features = selGenes$EnsID, scale =F,
        cluster.idents = T, dot.min = 0, dot.scale = 8, scale.by = "size") +
  scale_color_viridis_c() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=selGenes$EnsID, labels=selGenes$gene) +
  xlab("") + ylab("")



```


## vis BMP inhibitors {.tabset}
### violin
```{r vis BMP inhibitors violin}

genesDat <- data.frame(EnsID=rownames(seurat)) %>% 
  mutate(gene=gsub(".*\\.", "", EnsID))
selGenes <- data.frame(gene=c("Sost", "Sostdc1", "Chrd", "Ccn2", "Fst",
                              "Nog")) %>% 
  left_join(., genesDat, by="gene")

Idents(seurat) <- seurat$treatment
#seuratSub <- subset(seurat, downsample = min(table(seurat$cond2)))

pList <- sapply(selGenes$EnsID, function(x){
  p <- VlnPlot(object = seurat, features = x,
               group.by = "treatment",
               cols = colTreat, pt.size = 0
               )
  plot(p)
})

pList <- sapply(selGenes$EnsID, function(x){
  p <- VlnPlot(object = seurat, features = x,
               group.by = "label", split.by = "treatment",
               cols = colTreat, pt.size = 0
               )
  plot(p)
})


```


### heatmap
```{r vis BMP inhibitors heatmap, fig.height=6, fig.width=14}
selGenesHM <- selGenes %>% mutate(gene = EnsID)

Idents(seurat) <- seurat$label_plus_treat
gapVecCol <- seq(2, length(unique(seurat$label_plus_treat)), by=2)
pOut <- avgHeatmap(seurat = seurat, selGenes = selGenesHM,
                    colVecIdent = colLab, colVecCond=colTreat,
                    ordVec=levels(seurat$label_plus_treat),
                    gapVecR=NULL, gapVecC=gapVecCol,cc=FALSE,
                    cr=T, condCol=T)

```

### dotplot across label
```{r vis BMP inhibitors Dotplot across label, fig.height=5, fig.width=9}

Idents(seurat) <- seurat$label_plus_treat
DotPlot(seurat, assay="RNA", features = selGenes$EnsID, scale =T,
        cluster.idents = F, dot.min = 0, dot.scale = 3, scale.by = "size") +
  scale_color_viridis_c() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=selGenes$EnsID, labels=selGenes$gene) +
  xlab("") + ylab("")

DotPlot(seurat, assay="RNA", features = selGenes$EnsID, scale =F,
        cluster.idents = F, dot.min = 0, dot.scale = 5, scale.by = "size") +
  scale_color_viridis_c() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=selGenes$EnsID, labels=selGenes$gene) +
  xlab("") + ylab("")


```


### dotplot total
```{r vis BMP inhibitors Dotplot total, fig.height=4, fig.width=4}

Idents(seurat) <- seurat$treatment
DotPlot(seurat, assay="RNA", features = selGenes$EnsID, scale =F,
        cluster.idents = T, dot.min = 0, dot.scale = 8, scale.by = "size") +
  scale_color_viridis_c() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=selGenes$EnsID, labels=selGenes$gene) +
  xlab("") + ylab("")



```


## vis BMP downstream sign {.tabset}
### violin
```{r vis BMP downstream violin}

genesDat <- data.frame(EnsID=rownames(seurat)) %>% 
  mutate(gene=gsub(".*\\.", "", EnsID))
selGenes <- data.frame(gene=c("Id1", "Id2", "Id3", "Id4", "Ccn2", "Serpine1",
                              "Smad6", "Smad7","Bambi", "Rasef", "Msx1",
                              "Msx2", "Lef1", "Gata2", "Gata3", "Tbx2",
                              "Creb3l1", "Prdx2", "Klf10", "Snai1", "Bmpr2",
                              "Hey1", "Hes1")) %>% 
  left_join(., genesDat, by="gene")

Idents(seurat) <- seurat$treatment
#seuratSub <- subset(seurat, downsample = min(table(seurat$cond2)))

pList <- sapply(selGenes$EnsID, function(x){
  p <- VlnPlot(object = seurat, features = x,
               group.by = "treatment",
               cols = colTreat, pt.size = 0
               )
  plot(p)
})

pList <- sapply(selGenes$EnsID, function(x){
  p <- VlnPlot(object = seurat, features = x,
               group.by = "label", split.by = "treatment",
               cols = colTreat, pt.size = 0
               )
  plot(p)
})


```


### heatmap
```{r vis BMP downstream heatmap, fig.height=6, fig.width=14}
selGenesHM <- selGenes %>% mutate(gene = EnsID)

Idents(seurat) <- seurat$label_plus_treat
gapVecCol <- seq(2, length(unique(seurat$label_plus_treat)), by=2)
pOut <- avgHeatmap(seurat = seurat, selGenes = selGenesHM,
                    colVecIdent = colLab, colVecCond=colTreat,
                    ordVec=levels(seurat$label_plus_treat),
                    gapVecR=NULL, gapVecC=gapVecCol,cc=FALSE,
                    cr=T, condCol=T)

```

### dotplot across label
```{r vis BMP downstream Dotplot across label, fig.height=8, fig.width=9}

Idents(seurat) <- seurat$label_plus_treat
DotPlot(seurat, assay="RNA", features = selGenes$EnsID, scale =T,
        cluster.idents = F, dot.min = 0, dot.scale = 3, scale.by = "size") +
  scale_color_viridis_c() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=selGenes$EnsID, labels=selGenes$gene) +
  xlab("") + ylab("")

DotPlot(seurat, assay="RNA", features = selGenes$EnsID, scale =F,
        cluster.idents = F, dot.min = 0, dot.scale = 5, scale.by = "size") +
  scale_color_viridis_c() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=selGenes$EnsID, labels=selGenes$gene) +
  xlab("") + ylab("")


```


### dotplot total
```{r vis BMP downstream Dotplot total, fig.height=8, fig.width=4}

Idents(seurat) <- seurat$treatment
DotPlot(seurat, assay="RNA", features = selGenes$EnsID, scale =F,
        cluster.idents = T, dot.min = 0, dot.scale = 8, scale.by = "size") +
  scale_color_viridis_c() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=selGenes$EnsID, labels=selGenes$gene) +
  xlab("") + ylab("")



```




## session info
```{r session info}
sessionInfo()
date()
```




