---
title: "visualize total cells untreated"
author: "Mechthild Lütge"
date: "24 July 2020"
output:
  html_document:
    self_contained: no
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
set.seed(1000)
```

## load packages
```{r libs}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(tidyverse)
  library(Seurat)
  library(magrittr)
  library(dplyr)
  library(purrr)
  library(ggplot2)
  library(here)
  library(runSeurat3)
  library(ggsci)
  library(pheatmap)
  library(ggpubr)
})

```

## set dir
```{r set dir}
basedir <- here()
seurat <- readRDS(file = paste0(basedir,
                      "/data/MuHeart_totalCellsPlusTreated_integratedLabeled", 
                      "_seurat.rds"))
Idents(seurat) <- seurat$label
seurat$age[which(seurat$treatment != "no")] <- "12wk"

```



## color vectors
```{r colVec}

colPal <- pal_igv()(length(seurat$integrated_snn_res.0.4))
colGrp <- pal_jama()(length(unique(seurat$grp)))
colAge <- pal_uchicago()(length(unique(seurat$age)))
colStype <- pal_aaas()(length(unique(seurat$seqType)))
colDat <- c(pal_npg()(10),
            pal_futurama()(10))[1:(length(unique(seurat$dataset)))]
colTreat <- pal_jama()(length(unique(seurat$treatment)))

names(colPal) <- unique(seurat$integrated_snn_res.0.4)
names(colGrp) <- unique(seurat$grp)
names(colAge) <- unique(seurat$age)
names(colStype) <- unique(seurat$seqType)
names(colDat) <- unique(seurat$dataset)
names(colTreat) <- unique(seurat$treatment)

colLab <- c("#f4c37e", "#e73213", "#209ab9", "#df7f13",
            "#609eca", "#caa160", "#dbb6ca", "#98bfe8",
            "#7c0606", "#c2a17d", "#db27ab", "#dadada",
            "#32b4a5", "#4c8376")

names(colLab) <- c("InfMonos", "Tcells", "Endothelial", "Macrophages",
                   "Fibroblasts", "pDCsDCs", "NKcells", "Perivascular",
                   "Neutrophils", "Bcells", "Cardiomyocytes", "proliferating",
                   "endocardialEndothelial", "capEndothelial")


```


## vis data all {.tabset}
### clusters
```{r vis cluster all}

DimPlot(seurat, group.by ="integrated_snn_res.0.4", reduction = "umap",
        cols=colPal)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, group.by ="integrated_snn_res.0.4", reduction = "umap",
        cols=colPal)+
  theme_void()
```

### Label
```{r vis label all}

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab,
        shuffle = T)+
  theme_void() 

```

### group
```{r vis grp all}

DimPlot(seurat, reduction = "umap", group.by = "grp", cols=colGrp)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "grp", cols=colGrp, shuffle = T)+
  theme_void() 

```


### Sample
```{r vis sample all, fig.height=4, fig.width=7}

DimPlot(seurat, reduction = "umap", group.by = "dataset", cols=colDat)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "dataset", cols=colDat)+
  theme_void()

```


### Age
```{r vis age all}

DimPlot(seurat, reduction = "umap", group.by = "age", cols=colAge)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "age", cols=colAge, shuffle = T)+
  theme_void() 

```


### SeqType
```{r vis seqType all}

DimPlot(seurat, reduction = "umap", group.by = "seqType", cols=colStype)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "seqType", cols=colStype)+
  theme_void() 
```

### Treatment
```{r vis treat all}

DimPlot(seurat, reduction = "umap", group.by = "treatment", cols=colTreat)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "treatment", cols=colTreat,
        shuffle = T)+
  theme_void() 

```

### cnt Tab all
```{r cnts par lab per seqType}

table(seurat$label, seurat$seqType)
table(seurat$label, seurat$grp)
table(seurat$label, seurat$treatment)
```



## vis only untreated {.tabset}
```{r subset to only untreated}

seurat <- subset(seurat, treatment=="no")

```


### clusters
```{r vis cluster untreated}

DimPlot(seurat, group.by ="integrated_snn_res.0.4", reduction = "umap",
        cols=colPal)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, group.by ="integrated_snn_res.0.4", reduction = "umap",
        cols=colPal)+
  theme_void()
```

### Label
```{r vis label untreated}

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab,
        shuffle = T)+
  theme_void() 

```


### group
```{r vis grp untreated}

DimPlot(seurat, reduction = "umap", group.by = "grp", cols=colGrp)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "grp", cols=colGrp, shuffle = T)+
  theme_void() 

```


### Sample
```{r vis sample untreated, fig.height=4, fig.width=7}

DimPlot(seurat, reduction = "umap", group.by = "dataset", cols=colDat)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "dataset", cols=colDat)+
  theme_void()

```


### Age
```{r vis age untreated}

DimPlot(seurat, reduction = "umap", group.by = "age", cols=colAge)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "age", cols=colAge, shuffle = T)+
  theme_void() 

```


### SeqType
```{r vis seqType untreated}

DimPlot(seurat, reduction = "umap", group.by = "seqType", cols=colStype)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "seqType", cols=colStype)+
  theme_void() 
```

### Treatment
```{r vis treat untreated}

DimPlot(seurat, reduction = "umap", group.by = "treatment", cols=colTreat)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "treatment", cols=colTreat,
        shuffle = T)+
  theme_void() 

```

### Label split by cond
```{r vis label untreated split cond, fig.height=3.5, fig.width=8}

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab,
        split.by = "grp")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", group.by = "label", cols=colLab,
        split.by = "grp")+
  theme_void() 

```


### cnt Tab untreated
```{r cnts par lab untreat}

table(seurat$label, seurat$seqType)
table(seurat$label, seurat$grp)
```



## vis BMPs untreated {.tabset}
### Bmp2
```{r vis Bmp2 untreated}
p <- FeaturePlot(seurat, reduction = "umap", 
            features = "ENSMUSG00000027358.Bmp2",
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
p

```

### Bmp4
```{r vis Bmp4 untreated}
p <- FeaturePlot(seurat, reduction = "umap", 
            features = "ENSMUSG00000021835.Bmp4",
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
p

```

### Bmpr1a
```{r vis Bmpr1a untreated}
p <- FeaturePlot(seurat, reduction = "umap", 
            features = "ENSMUSG00000021796.Bmpr1a",
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
p

```

### Bmpr2
```{r vis Bmpr2 untreated}
p <- FeaturePlot(seurat, reduction = "umap", 
            features = "ENSMUSG00000067336.Bmpr2",
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
p

```


## vis BMPs untreated split by cond {.tabset}
### Bmp2
```{r vis Bmp2 untreated split, fig.height=4, fig.width=9}
p <- FeaturePlot(seurat, reduction = "umap", split.by = "grp",
            features = "ENSMUSG00000027358.Bmp2",
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
p

```

### Bmp4
```{r vis Bmp4 untreated split, fig.height=4, fig.width=9}
p <- FeaturePlot(seurat, reduction = "umap", split.by = "grp",
            features = "ENSMUSG00000021835.Bmp4",
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
p

```

### Bmpr1a
```{r vis Bmpr1a untreated split, fig.height=4, fig.width=9}
p <- FeaturePlot(seurat, reduction = "umap", split.by = "grp", 
            features = "ENSMUSG00000021796.Bmpr1a",
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
p

```

### Bmpr2
```{r vis Bmpr2 untreated split, fig.height=4, fig.width=9}
p <- FeaturePlot(seurat, reduction = "umap", split.by = "grp",
            features = "ENSMUSG00000067336.Bmpr2",
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
p

```


## save object untreated with label
```{r save int untreated}
saveRDS(seurat, file = paste0(basedir,
                      "/data/MuHeart_totalCells_integratedLabeled_seurat.rds"))

```


## session info
```{r session info}
sessionInfo()
date()
```




